Electric polarizability as the fundamental variable in the dielectric properties of two-dimensional materials

### 2019-01-04 SI.tex
Fix minor errors in the manuscript
Implement changes of SI from elton 20181224

\item Supplementary Section \ref{sec:2D-3D-rescale}: Hypothetical 2D ``dielectric constant'' rescaled from 3D dielectric constant

\item Supplementary Section \ref{sec:theory-1}: Polarizability-Based Theoretical Model

@@

\item Supplementary Section \ref{sec:gpaw}: Validation on the C2DB 2D materials database

@@

In this section we provide more analysis on the dielectric properties
of 2D materials, with focus on the dielectric properties calculated
using many-body Green function method (G$_0$W$_0$) and the
frequency-dependency of the superlattice dielectric functions and 2D
polarizabilities.

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta \rho=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}. The
density $\rho$ is calculated as the average value in the
\textit{xy}-plane, and the value is shown in Figure
\ref{fig:rho-profile}. As can be seen the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{}. When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus giving converged values of $\alpha_{\mathrm{2D}}$.

@@

Here we will show the results of dielectric properties calculated
using higher level of many-body Green function method (G$_0$W$_{0}$)
and with electron-hole interaction. We first compare the case of a
layer of hBN within an superlattice with varying size with the
dielectric properties calculated using PBE and G$_{0}$W$_{0}$ methods,
as shown in Figure
\ref{fig:GW-PBE-alpha}. $\varepsilon_{\mathrm{SL}}^{\parallel}$ and
$\varepsilon_{\mathrm{SL}}^{\perp}$ as functions of $L$, shown in
Figure \ref{fig:GW-PBE-alpha}a and \ref{fig:GW-PBE-alpha}b,
respectively, do not converge with $L$, while the corresponding 2D
polarizabilities in Figure \ref{fig:GW-PBE-alpha}c and
\ref{fig:GW-PBE-alpha}d are almost $L$-independent. It is worthnoting
that since the G$_{0}$W$_{0}$ method has better estimation of the bandgap, $\varepsilon_{\mathrm{SL}}$ and $\alpha_{\mathrm{2D}}$ are smaller using G$_{0}$W$_{0}$ than PBE method.
\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-alpha-zero.pdf}
 \caption{Variation of dielectric properties and
   $\varepsilon^{\perp}_{\mathrm{SL}}$ as a function of $L$ for
   monolayer BN calculated at the level of PBE and G$_{\rm 0}$W$_{0}$.
   \textbf{a}. $\varepsilon_{\mathrm{SL}}^{\parallel}$ as function of
   $L$. \textbf{a}. $\alpha_{\mathrm{2D}}^{\parallel}$ as function of
   $L$ corresponding to
   \textbf{b}. \textbf{c}. $\varepsilon_{\mathrm{SL}}^{\perp}$ as
   function of $L$.  \textbf{d}. $\alpha_{\mathrm{2D}}^{\perp}$ as
   function of $L$. In all cases the values obtained using
   G$_{0}$W$_{0}$ method is smaller than that using PBE method due to
   better estimation of the bandgap.}
  \label{fig:GW-PBE-alpha}
\end{figure}

@@

We calculate the frequency-dependent dielectric properties of hBN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$
with electron-hole interaction (G$_{0}$W$_{0}$+BSE). As shown in
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out}, the
intensities of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range. The $L$-dependency can also be
removed using Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, yielding lattice-independent
polarizabilities throughout the frequency domain. Note when extracting
the frequency-dependent 2D polarizabilities from
Eqs. \ref{main-eq:alpha-para-def} and \ref{main-eq:alpha-perp-def},
the peak position from $\varepsilon_{\mathrm{SL}}$ is preserved. This
is explained by the fact that the local extrema from spectra of
$\varepsilon_{\mathrm{SL}}$ is also the local extrema in corresponding
$\alpha_{\mathrm{2D}}$, since when
$\partial \varepsilon_{\mathrm{SL}}(\omega) / \partial \omega = 0 $,
we have
\begin{subequations}
\begin{eqnarray}
  \label{eq:extrema-para}
  \frac{\partial \alpha_{\mathrm{2D}}^{\parallel}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0   \\
  \label{eq:extrema-perp}
  \frac{\partial \alpha_{\mathrm{2D}}^{\perp}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{1}{\varepsilon_{\mathrm{SL}}^{2}}\frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0
\end{eqnarray}
\end{subequations}
which indicates that the peak position is perserved when transforming
$\varepsilon_{\mathrm{SL}}(\omega)$ to $\alpha_{\mathrm{2D}}(\omega)$.

@@

\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-in_plane.pdf}
 \caption{Dependence of in-plane dielectric properties on $L$ for
   monolayer BN calculated using PBE method: \textbf{a} Real part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{b} Real part of
   $\alpha_{\mathrm{2D}}^{\parallel}$, \textbf{c} Imaginary part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{d} Imaginary part
   of $\alpha_{\mathrm{2D}}^{\parallel}$.  The same {\bf k}-sampling
   is used in all simulations with the only variable quantity being
   $L$. Clearly, the $L$-dependency of the superlattice dielectric
   function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.9\linewidth]{img/SI-PBE-omega-out_of_plane.pdf}
  \caption{Dependence of out-of-plane dielectric properties on $L$ for
    monolayer BN calculated using PBE method: \textbf{a} Real part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{b} Real part of
    $\alpha_{\mathrm{2D}}^{\perp}$, \textbf{c} Imaginary part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{d} Imaginary part of
    $\alpha_{\mathrm{2D}}^{\perp}$.  The same {\bf k}-sampling is used
    in all simulations with the only variable quantity being
    $L$. Clearly, the $L$-dependency of the superlattice dielectric
    function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-out}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-in_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-in} but calculated
   using G$_{0}$W$_{0}$ method. A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-omega-out_of_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-out} but calculated
   using G$_{0}$W$_{0}$ method.  A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-out}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-in_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-in} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE).  The peak shift
   in energies are reduced in comparison with G$_{\rm 0}$W$_{0}$. As a
   result the polarizabilities becomes almost independent of $L$,
   consistent to results obtained by the PBE method.}
  \label{fig:BSE-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-BSE-out_of_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-out} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ - BSE).  As a result
   the polarizabilities becomes almost independent of $L$, consistent
   to results obtained by the PBE method.  }
  \label{fig:BSE-omega-out}
\end{figure}

@@

%%%%%%\section{Theoretical Analysis}
%%%%%%\label{sec:theory}
\section{Polarizability-Based Theoretical Model}
\label{sec:theory-1}
The universal relations for $\alpha_{\mathrm{2D}}^{\parallel}$ and
$\alpha_{\mathrm{2D}}^{\perp}$ revealed by main text
Eqs. \ref{main-eq:2D-Moss-para} and \ref{main-eq:2D-Moss-perp} are not
coincidence. Combining recent theoretical findings of the linear
relation between exciton binding energy $E_{\mathrm{b}}$ and
$E_{\mathrm{g}}$ of 2D materials
\cite{Choi_linear_2015,Olsen_2016_hydrogen,Jiang_2017_Eg_Eb}, and the
fact that the $E_{\mathrm{b}}$ is roughly inversely proportional to
$\alpha_{\mathrm{2D}}^{\parallel}$ \cite{Pulci_2014}, it is reasonable
to have a general Moss-like relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{-1}$. Moreover,
the bandgap-independent relation of 2D $\alpha_{\mathrm{2D}}^{\perp}$
resembles molecular polarizabilities of conjugate molecules
\cite{Davies_1952}, fullerenes \cite{Sabirov_2014} and carbon
nanotubes \cite{Benedict_1995}, which are also shown to be
geometry-dependent.

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wave function of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components, similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$, such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, c, v are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

@@

\subsection{Detailed derivations for main text Eq. \ref{main-eq:2D-Moss-para}}  %
\label{ssec:theory-1-para}

In this section we show how main text Eq. \ref{main-eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Following the method of $\mathbf{k} \cdot \mathbf{p}$ theory from
Ref. \citenum{Jiang_2017_Eg_Eb}, the matrix element in the numerator
of Eq. \ref{eq:alpha_para_RPA} is approximated by:

\begin{equation}
\label{eq:matrix-approx}
|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}
\approx {\displaystyle \frac{\hbar^{2}}{2 m^{*}}
  \frac{1}{E_{\mathrm{g}} + \frac{\hbar^{2} k^{2}}{2 m^{*}}}}
\end{equation}
plug it into Eq. \ref{eq:alpha_para_RPA} and integrate within the 2D Brillouin Zone from $|k|=0$ to $|k|= k_{\mathrm{BZ}}$, where $k_{\mathrm{BZ}}$ is the wavevector at the boundary of the 2D Brillouin Zone, we get:
\begin{equation}
  \label{eq:derivation-2D-Moss-para}
  \begin{aligned}
\alpha_{\mathrm{2D}}^{\parallel} &= N\cdot -\frac{e^{2}}{2 \pi}
\frac{1}{E_{g} + \beta} \Biggr\vert_{\beta=0}^{\beta=\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}}} \\
&\approx N e^{2}/(2\pi E_{\mathrm{g}}) = C^{\parallel} E_{g}^{-1}
\end{aligned}
\end{equation}
where $N$ is de degeneracy of bands associated with $E_{g}$. The
approximation in Eq. \ref{eq:derivation-2D-Moss-para} is due to the
fact that $\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}} \gg E_{g}$, and
we arrive at main text Eq. \ref{main-eq:2D-Moss-para}.

>>>>

\item Supplementary Section \ref{sec:2D-3D-rescale}: Hypothetical 2D ``dielectric constant'' rescaled from the 3D dielectric constant

\item Supplementary Section \ref{sec:theory-1}: Polarizability-based theoretical model

@@

\item Supplementary Section \ref{sec:gpaw}: Validation of results using a larger 2D material database

@@

In this section we provide more analysis on the dielectric properties
of 2D materials calculated using many-body Green function method 
(G$_0$W$_0$), including electron-hole interactions at the level of the 
Bethe-Salpeter equation (G$_0$W$_0$--BSE), and at the frequency-dependent regime.

\subsection{Profile of induced dipoles of 2D material}
\label{sec:dipole-plot}
Here we show in detail the
$\Delta {\rho}=\rho(\boldsymbol{E}) - \rho(\boldsymbol{E}=0)$ profile of
the 2H-MoS$_{2}$ slab in main text Figure \ref{main-fig-1}. The
density $\Delta \rho$ is calculated via $\Delta \rho(z) = \frac{1}{S} \int_{S} \Delta \rho (x,y;z) dx dy $, 
where $S$ is the surface of the unit cell perpendicular to the given direction, in this case $z$. 
%
As can be seen in Figure \ref{fig:rho-profile} the induced charges on the
MoS$_{2}$ layer only extends to a width of $\sim{}$12 \AA{} centered at the middle of the layer. This corresponds to about 6 \AA ~from each side. 
When the
SL size $L \gg$12 \AA{} as in the first principle calculations done in
the main text, the induced dipoles from the periodic images do not
interact, and thus give the converged values of $\alpha_{\mathrm{2D}}$.

@@

Here we will show the results of dielectric properties calculated
using many-body Green function method (G$_0$W$_{0}$)
and with electron-hole interactions at the level of the Bethe-Salpeter 
equation (G$_0$W$_{0}$--BSE). We first compare the case of a
monolayer of BN within a varying superlattice $L$ 
calculated using PBE and G$_{0}$W$_{0}$ as shown in Figure \ref{fig:GW-PBE-alpha}. 
%
%
Both $\varepsilon_{\mathrm{SL}}^{\parallel}$ and
$\varepsilon_{\mathrm{SL}}^{\perp}$ do not converge as a function of $L$ 
despite of the separation utilized in the simulations (Figure \ref{fig:GW-PBE-alpha}{\bf a}-{\bf c}). However, corresponding 2D polarizabilities are almost $L$-independent (Figure \ref{fig:GW-PBE-alpha}{\bf b}-{\bf d}). It is worth noting
that since the G$_{0}$W$_{0}$ method has better estimation of the bandgap, $\varepsilon_{\mathrm{SL}}$ and $\alpha_{\mathrm{2D}}$ are smaller using G$_{0}$W$_{0}$ than in PBE functionals.
\begin{figure}[htbp]
  \centering
 \includegraphics[width=0.9\linewidth]{img/SI-GW-alpha-zero.pdf}
 \caption{{\bf Variation of 
   $\varepsilon^{\perp}_{\mathrm{SL}}$ as a function of $L$ for
   monolayer BN calculated at the level of PBE and G$_{\rm 0}$W$_{0}$.}
   \textbf{a}-{\bf b}. $\varepsilon_{\mathrm{SL}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\parallel}$ as function of $L$, respectively.    
  \textbf{c}-{\bf d}. $\varepsilon_{\mathrm{SL}}^{\perp}$ and 
  $\alpha_{\mathrm{2D}}^{\perp}$ as
   function of $L$. 
   In all cases the values obtained using
   G$_{0}$W$_{0}$ method is smaller than that using PBE method due to
   better estimation of the bandgap.}
  \label{fig:GW-PBE-alpha}
\end{figure}

@@

We calculate the frequency-dependent dielectric properties of BN with
varying supperlattice $L$ at PBE, G$_{0}$W$_{0}$, and G$_{0}$W$_{0}$--BSE. 
Figure \ref{fig:PBE-omega-in} and \ref{fig:PBE-omega-out} show the
magnitudes of $\varepsilon_{\mathrm{SL}}^{\parallel}(\omega)$ and
$\varepsilon_{\mathrm{SL}}^{\perp}(\omega)$ reduce with increasing $L$
throughout the whole frequency range at PBE level. The $L$-dependency can also be
removed using Eqs. \ref{main-eq:alpha-para-def} and
\ref{main-eq:alpha-perp-def}, yielding lattice-independent
polarizabilities throughout the frequency domain (Figure \ref{fig:PBE-omega-in}{\bf b}-{\bf d} and \ref{fig:PBE-omega-out}{\bf b}-{\bf d}). 
%
Note that when extracting
the frequency-dependent 2D polarizabilities from
Eqs. \ref{main-eq:alpha-para-def} and \ref{main-eq:alpha-perp-def},
the peak position in energy $\omega (eV)$ from $\varepsilon_{\mathrm{SL}}$ is preserved. This
is explained by the fact that the local extrema from spectra of
$\varepsilon_{\mathrm{SL}}$ is also the local extrema in corresponding
$\alpha_{\mathrm{2D}}$, since when
$\partial \varepsilon_{\mathrm{SL}}(\omega) / \partial \omega = 0 $,
we have
\begin{subequations}
\begin{eqnarray}
  \label{eq:extrema-para}
  \frac{\partial \alpha_{\mathrm{2D}}^{\parallel}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0   \\
  \label{eq:extrema-perp}
  \frac{\partial \alpha_{\mathrm{2D}}^{\perp}(\omega)}{\partial \omega}
  &= \varepsilon_{0}L {\displaystyle \frac{1}{\varepsilon_{\mathrm{SL}}^{2}}\frac{\partial \varepsilon_{\mathrm{SL}}^{\parallel}}{\partial \omega}} = 0
\end{eqnarray}
\end{subequations}
which indicates that no corrections in energy are present 
when transforming
$\varepsilon_{\mathrm{SL}}(\omega)$ to $\alpha_{\mathrm{2D}}(\omega)$.

@@

\begin{figure}[htbp]
  \centering
 \includegraphics[width=1.00\linewidth]{img/SI-PBE-omega-in_plane.pdf}
 \caption{Dependence of in-plane dielectric properties on $L$ for
   monolayer BN calculated using PBE method: \textbf{a} Real part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{b} Real part of
   $\alpha_{\mathrm{2D}}^{\parallel}$, \textbf{c} Imaginary part of
   $\varepsilon_{\mathrm{SL}}^{\parallel}$, \textbf{d} Imaginary part
   of $\alpha_{\mathrm{2D}}^{\parallel}$.  The same {\bf k}-sampling
   is used in all simulations with the only variable quantity being
   $L$. Clearly, the $L$-dependency of the superlattice dielectric
   function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
  \includegraphics[width=1.0\linewidth]{img/SI-PBE-omega-out_of_plane.pdf}
  \caption{Dependence of out-of-plane dielectric properties on $L$ for
    monolayer BN calculated using PBE method: \textbf{a} Real part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{b} Real part of
    $\alpha_{\mathrm{2D}}^{\perp}$, \textbf{c} Imaginary part of
    $\varepsilon_{\mathrm{SL}}^{\perp}$, \textbf{d} Imaginary part of
    $\alpha_{\mathrm{2D}}^{\perp}$.  The same {\bf k}-sampling is used
    in all simulations with the only variable quantity being
    $L$. Clearly, the $L$-dependency of the superlattice dielectric
    function is removed using 2D polarizabilities.}
  \label{fig:PBE-omega-out}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=1.0\linewidth]{img/SI-GW-omega-in_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-in} but calculated
   using G$_{0}$W$_{0}$ method. A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=1.0\linewidth]{img/SI-GW-omega-out_of_plane.pdf}
 \caption{Similar to Figure \ref{fig:PBE-omega-out} but calculated
   using G$_{0}$W$_{0}$ method.  A shift of peak position (frequency)
   in the dielectric function spectra is observed, resulted from the
   change of quasi-particle energy in G$_{0}$W$_{0}$ calculations
   invloving the 2D slab.}
  \label{fig:GW-omega-out}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=1.0\linewidth]{img/SI-BSE-in_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-in} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ + BSE).  The peak shift
   in energies are reduced in comparison with G$_{\rm 0}$W$_{0}$. As a
   result the polarizabilities becomes almost independent of $L$,
   consistent to results obtained by the PBE method.}
  \label{fig:BSE-omega-in}
\end{figure}

\begin{figure}[htbp]
  \centering
 \includegraphics[width=1.0\linewidth]{img/SI-BSE-out_of_plane.pdf}
 \caption{Similar as in Fig. \ref{fig:GW-omega-out} but taking into
   account electron-hole interactions at the level of the
   Bethe-Salpeter equation (G$_{\rm 0}$W$_{0}$ + BSE).  As a result
   the polarizabilities becomes almost independent of $L$, consistent
   to results obtained by the PBE method.}
  \label{fig:BSE-omega-out}
\end{figure}

@@

%%%%%%\section{Theoretical Analysis}
%%%%%%\label{sec:theory}
\section{Polarizability-Based Theoretical Model}
\label{sec:theory-1}
The universal relations for $\alpha_{\mathrm{2D}}^{\parallel}$ and
$\alpha_{\mathrm{2D}}^{\perp}$ revealed by main text
Eqs. \ref{main-eq:2D-Moss-para} and \ref{main-eq:2D-Moss-perp} are not
coincidental. Combining recent theoretical findings of the linear
relation between exciton binding energy $E_{\mathrm{b}}$ and
$E_{\mathrm{g}}$ of 2D materials
\cite{Choi_linear_2015,Olsen_2016_hydrogen,Jiang_2017_Eg_Eb}, and the
fact that the $E_{\mathrm{b}}$ is roughly inversely proportional to
$\alpha_{\mathrm{2D}}^{\parallel}$ \cite{Pulci_2014}, it is reasonable
to have a general relation between
$\alpha_{\mathrm{2D}}^{\parallel}$ and $E_{\mathrm{g}}^{-1}$. Moreover,
the bandgap-independent relation of 2D $\alpha_{\mathrm{2D}}^{\perp}$
resembles molecular polarizabilities of conjugate molecules
\cite{Davies_1952}, fullerenes \cite{Sabirov_2014} and carbon
nanotubes \cite{Benedict_1995}, which are also shown to be
geometry-dependent.

In this section we show in detail the polarizability-based theoretical
framework that leads to the 2D Moss-like relations proposed in the
main text.  Due to its highly anisotropic nature, the wave function of
an isolated 2D material $\psi(\mathbf{r})$ can be separated into the
in- and out-of-plane components ($\psi^{\parallel}(\boldsymbol{\rho})$ and $\psi^{\perp}(z)$) 
similar to the treatment of quantum
wells (QW),\cite{davies_physics_1997} such that
$\psi=\psi^{\parallel}\psi^{\perp}$, where $\boldsymbol{\rho}=(x, y)$
is the in-plane coordinate. Using the Bloch theorem, the periodic
$\psi^{\parallel}(\boldsymbol{\rho})$ can be further expressed as
$\psi^{\parallel}(\boldsymbol{\rho})=e^{i\mathbf{k} \cdot
  \boldsymbol{\rho}}u(\boldsymbol{\rho})$, where $\mathbf{k}$ is the
in-plane wave vector and $u(\boldsymbol{\rho})$ is periodic function
in the xy-plane. According to the random phase approximation (RPA)
theory\cite{Adler_1962}, $\varepsilon_{\mathrm{SL}}$ is the
$\mathbf{q} \to 0$ and $\omega \to 0$ limits of the non-interacting
dielectric function $\varepsilon(\mathbf{q}, \omega)$, where
$\mathbf{q}$ is the momentum transfer and $\omega$ is the frequency:
\begin{equation}
  \label{eq:RPA-eps2}
  \varepsilon_{\mathrm{SL}}
  = \lim_{\mathbf{q} \to 0} 1 + \frac{2e^{2}}{\varepsilon_{0} |\mathbf{q}|^{2} \Omega}
  \sum_{\mathrm{k, c, v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
  \left[f(\psi_{\mathrm{c}}) - f(\psi_{\mathrm{v}})\right]
\end{equation}
where $e$ is the unit charge, $c$, $v$ are the conduction and valence
bands, $E$ is the eigenenergy of individual bands, and $f$ is the
Fermi-Dirac distribution function. Taking the
limit that $L\to\infty$, when
$\varepsilon^{\perp}_{\mathrm{SL}} \approx 1$, we have
$1-1/\varepsilon^{\perp}_{\mathrm{SL}} \approx
(\varepsilon_{\mathrm{SL}}^{\perp} - 1)$. Therefore
$\alpha_{\mathrm{2D}}^{\parallel}$ and $\alpha_{\mathrm{2D}}^{\perp}$
at 0 K can be unified by the same equation:
\begin{equation}
  \label{eq:alpha-RPA}
  \alpha = \frac{2e^{2}}{|q|^{2}A} \sum_{\mathrm{k,c,v}}
  \frac{|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k+q}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
where the direction is determined by $\mathbf{q}$. Next we will show
that the different behavior of $\psi^{\parallel}$ and $\psi^{\perp}$
give rise to the main text Eqs. \ref{main-eq:2D-Moss-para} and
\ref{main-eq:2D-Moss-perp}.

@@

\subsection{Detailed derivations of Eq. \ref{main-eq:2D-Moss-para}}  %
\label{ssec:theory-1-para}

In this section we show how Eq. \ref{main-eq:alpha-para-def}
is derived from Eq. \ref{eq:alpha-RPA}. For the in-plane component
$\alpha_{\mathrm{2D}}^{\parallel}$, $e^{-i\mathbf{qr}}$ is independent
of $z$, therefore the integral in
$|<\psi_{\mathrm{v}}(\mathbf{k})|e^{-i\mathbf{q}\mathbf{r}}|\psi_{\mathrm{c}}(\mathbf{k+q})>|^{2}$
becomes independent of $\psi^{\perp}$, due to the orthogonality
and normality. The Bloch-wave form of $\psi^{\parallel}$ ensures that
only the cell-function $u(\mathbf{k})$ contributes to the final result
of $\alpha^{\parallel}_{\mathrm{2D}}$ \cite{davies_physics_1997}, such
that:
\begin{equation}
  \label{eq:alpha_para_RPA}
  \alpha_{\mathrm{2D}}^{\parallel} = \frac{2e^{2}}
  {(2 \pi)^{2}} \int d^{2}\mathbf{k} \sum_{\mathrm{c, v}}
  \frac{|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}}
  {E_{\mathrm{c}}(\mathbf{k}) - E_{\mathrm{v}}(\mathbf{k})}
\end{equation}
Following the method of $\mathbf{k} \cdot \mathbf{p}$ theory from
Ref. \citenum{Jiang_2017_Eg_Eb}, the matrix element in the numerator
of Eq. \ref{eq:alpha_para_RPA} is approximated by:

\begin{equation}
\label{eq:matrix-approx}
|<u_{\mathrm{c}}(\mathbf{k})|\nabla|u_{\mathrm{v}}(\mathbf{k})>|^{2}
\approx {\displaystyle \frac{\hbar^{2}}{2 m^{*}}
  \frac{1}{E_{\mathrm{g}} + \frac{\hbar^{2} k^{2}}{2 m^{*}}}}
\end{equation}
plug it into Eq. \ref{eq:alpha_para_RPA} and integrate within the 2D Brillouin zone from $|k|=0$ to $|k|= k_{\mathrm{BZ}}$, where $k_{\mathrm{BZ}}$ is the wavevector at the boundary of the 2D Brillouin Zone, we get:
\begin{equation}
  \label{eq:derivation-2D-Moss-para}
  \begin{aligned}
\alpha_{\mathrm{2D}}^{\parallel} &= N\cdot -\frac{e^{2}}{2 \pi}
\frac{1}{E_{g} + \beta} \Biggr\vert_{\beta=0}^{\beta=\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}}} \\
&\approx N e^{2}/(2\pi E_{\mathrm{g}}) = C^{\parallel} E_{g}^{-1}
\end{aligned}
\end{equation}
where $N$ is de degeneracy of bands associated with $E_{g}$. The
approximation in Eq. \ref{eq:derivation-2D-Moss-para} is due to the
fact that $\frac{\hbar^{2} k^{2}_{\mathrm{BZ}}}{m^{*}} \gg E_{g}$, and
we arrive at Eq. \ref{main-eq:2D-Moss-para}.
